---
title: "IMAGE ANALYSIS FOR LESION DETECTION - BY KLIEBENSTEIN LAB"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---
# Description of the pipeline

1. The images are cropped to keep only the necessary part of the tray containing the leaves. Removing unnecessary part of images make the image analysis faster and more reliable. The cropped images are imported into R and transformed into hue/saturation/value (hsv) color space.

2. For each image, the leaf mask is created by selecting all pixels above Otsu’s saturation threshold and below Otsu’s value threshold, opening using a 15-pixel disc-shaped kernel

3. A mask marking the lesion within the leaf is created by excluding all pixels that had a green or yellow hue within the leaf mask, opening using a 10-pixel disc-shaped kernel. 

4. Masks are visually confirmed by overlaying the masks with the original image in ImageJ (https://imagej.net/Welcome).

5. All leaves and lesions are phenotyped using the masks and the original raw image using the computeFeatures() function in ‘EBImage’ which measures ~900 characteristics per object including spatial characteristics (area, perimeter, major and minor axes, and eccentricity) and texture characteristics (Haralick texture transformations) in pixels. Leaf and lesion objects were also measured for 25 custom and composite phenotypes including the number of green and yellow pixels the proportion of yellow to green pixels or residual leaf area and the proportion of lesion area to leaf area. 


## R packages required

This pipeline is constructed around the packages EBImage and CRImage available on Bioconductor. For more information:

EBImage: https://www.bioconductor.org/packages/release/bioc/html/EBImage.html

CRImage: https://bioconductor.org/packages/release/bioc/html/CRImage.html

```{r}
# To install, run:
#source("http://bioconductor.org/biocLite.R")
#biocLite("EBImage")
#biocLite("CRImage")

# To use, run:
library("EBImage")
library("CRImage")
```

## Creating the leaf and lesion masks

```{r}

# Enter your work directory below:
#setwd("/.../...../")

#note:EBImage can read jpeg, png and tiff. 

Files <- list.files(pattern=".jpg")
Files

for(i in 1:length(Files)) {
#tracks time to run script
print(Sys.time())

#reads image into a file for EBImage
Plate.cr2 <- readImage(Files[i])

#retrieves dimensions of image file
Plate.row <- dim(Plate.cr2)[1]
Plate.col <- dim(Plate.cr2)[2]

# Import Image, get dimensions, and separate channels
Plate.red <- as.vector(Plate.cr2[,,1])
Plate.grn <- as.vector(Plate.cr2[,,2])
Plate.blu <- as.vector(Plate.cr2[,,3])

# Convert Image from rgb to hsv 
# The 14-bit pixel value range is 0:16383  ]
Plate.hsv <- rgb2hsv(Plate.red, Plate.grn, Plate.blu, maxColorValue = 16383)
Plate.hue <- Image(Plate.hsv[1,], dim = c(Plate.row, Plate.col))
Plate.sat <- Image(Plate.hsv[2,], dim = c(Plate.row, Plate.col))
Plate.value <- Image(Plate.hsv[3,]*10000, dim = c(Plate.row, Plate.col))
rm(Plate.red, Plate.grn, Plate.blu, Plate.hsv)

#fills holes in objects
Lf.mask <- fillHull(Plate.hue > 0.1 & Plate.hue < 0.3)
#sigma sets SD of gaussian shape
kern <- makeBrush(7,shape = 'gaussian',sigma = 0.3)
Lf.mask <- opening(Lf.mask,kern)

#adds labels to objects
Obj <- bwlabel(Lf.mask)
Obj.minRad <- computeFeatures.shape(Obj)[,'s.radius.min']
Obj.lvs <- as.numeric(names(Obj.minRad[Obj.minRad > 5]))
Lf.mask <- Image(Obj %in% Obj.lvs, dim = c(Plate.row, Plate.col))

writeImage(Lf.mask, gsub(".jpg", "_LeafMask.jpg", Files[i]))

# Identify lesions
# Try modifications here!
Lesion.mask <- fillHull((Plate.sat<0.38 * Lf.mask))
Lesion.mask <- fillHull((Plate.sat>0.2 & Plate.sat<0.5 * Lf.mask))

### Fill Holes, Filter within Leaves and Open 
# Lesion.mask <- fillHull(Lesion.mask)
# Lesion.mask <- Lesion.mask & Leaf.Area.Mask
kern2 <- makeBrush(3,shape='disc') #changed from 8: next odd number
Lesion.mask <- opening(Lesion.mask,kern2)
writeImage(Lesion.mask, gsub(".jpg", "_LesionMask.jpg", Files[i]))

}
```

## Mask editing tips

Masks are edited in ImageJ. To overlay the original image and the masks, open the mask and overlay the original image at 70% opacity. 

Use pencil tool with Alt button to draw black over pixels incorrectly labeled as leaf. 

The following can produce error messages in the next steps of the pipeline if not corrected:

- The lesion can not be larger than the leaf. The lesion mask area has to be within the leaf mask area.

- No other object than leaves should be kept as white dots or area (remove the remaining noise if necessary).

- No objects should be touching. For any objects that are touching, draw a >=1 pixel line between them in black.

- No leaves should touch the sides of the images (crop the images accordingly to begin with).

To save the masks, go to Image > Overlay > Remove Overlay. Save the black and white mask as tiff. The mask names should end as LeafMask.tif & LesionMask.tif.

The mask editing can be done in other image analysis software with the following warnings:

- The final masks can only be black and white pixels with direct transition. The masks can not contain any grey pixels. 

- In GIMP, the pencil tool with hardness at 1, opacity at 100%, aspect ratio at 0 can be used with caution. 

## Lesion and leaf caracterisation

Charge the following function
```{r message=FALSE, warning=FALSE}
# Function for labeling the images
AddLabelOnPicture <- function(orig.label , image.f = RawFiles[f]){
  
  # generate label image
  MakeLabelMarker <- function(){
    jpeg(filename = (gsub(".jpg","_labels.jpg",RawFiles[f])),  width = 1000 , height = 1000 ,  units = "px")
    #par.default <- par()
    
    par(mfrow=c(20,10),oma = c(0,0,0,0),mar = c(0,0,0,0) , bg ="white" )
    for( i in 0:199 ){
      plot.new()
      plot.window(xlim=c(0,1), ylim=c(0,1)) 
      #plot(0:1,0:1,type="n")
      text(0.5,0.5,i,col='black' , cex= 6)
    }
    dev.off()
    #par(par.default) 
    readImage(gsub(".jpg","_labels.jpg",RawFiles[f]))
  }
  
  labelMarker <- MakeLabelMarker()
  
  getLabelMarker <- function(i){
    x <- i%%10
    xa <- (x * 100 +1) : ((x+1) * 100)
    y <- (i-x)/10
    ya <- (y * 50 +1) : ((y+1) * 50)
    labelMarker[xa,ya,]
  }
  
  RawImage <-readImage(image.f)
  
  # start to draw labels
  for( i in 1:dim(orig.label)[1]) {
    row <- orig.label[i,]
    x <- round(row$Lesion.0.m.cx)
    y <- round(row$Lesion.0.m.cy)
    labelName <- as.numeric(as.character(row$Object))
    print(paste("DEBUG" , Sys.time() , ": row ",i,",  label " , labelName , ", Object " , row$Object) )
    label <- getLabelMarker(labelName)
    
    xa <- (x - 50 + 1 ) : (x + 50 )
    ya <- (y - 50 + 1 ) : (y )
    
    RawImage[xa,ya,] <- (label[,,1:3] + RawImage[xa,ya,]) # C code would be better
  }
  
  print(paste("DEBUG" , Sys.time() , ": output image " , gsub(".jpg","_labeled.jpg",RawFiles[f])) )
  writeImage(RawImage, gsub(".jpg","_labeled.jpg",RawFiles[f]))
}

```

The analysis of leaf and lesion is done with the following code.Depending on the characteristics of the computer used, this can take >10 minutes for each image. 

```{r}

setwd("/../../")

RawFiles <- list.files(pattern = ".jpg")
LeafMaskFiles <- list.files(pattern= "LeafMask.tiff")
LesionMaskFiles <- list.files(pattern="LesionMask.tiff")

if(length(RawFiles) == length(LeafMaskFiles) & length(RawFiles) == length(LesionMaskFiles)) {
  for(f in 1:length(RawFiles)) {
    
    print(Sys.time())
    print(RawFiles[f])
    print("Reading Files....")
    RawImage <- readImage(RawFiles[f])
    LeafMask <- readImage(LeafMaskFiles[f])
       # Run the following line if error in image dimension
       # LesionMask <- LesionMask[, ,1]
    LeafMask[LeafMask > 0.5] <- 1
    LeafMask[LeafMask < 0.5] <- 0
    LesionMask <- readImage(LesionMaskFiles[f])
      # Run the following line if error in image dimension
      # LesionMask <- LesionMask[, ,1]
    LesionMask[LesionMask > 0.5] <- 1
    LesionMask[LesionMask < 0.5] <- 0
    LeafResidual <- LeafMask & !LesionMask
    
    Image.row <- dim(RawImage)[1]
    Image.col <- dim(RawImage)[2]
    
    print("Calculating hue....")
    ImageHSV <- rgb2hsv(as.vector(RawImage[,,1]), as.vector(RawImage[,,2]), as.vector(RawImage[,,3]), maxColorValue = 16383)
    Image.hue <- Image(ImageHSV[1,], dim = c(Image.row, Image.col))
    rm(ImageHSV)
    
    #Yellow and green pix
    print("Identifying Yellow and Green pixels....")
    Yellow.pix <- Image.hue > 0.145 & Image.hue < 0.188 & LeafMask > 0.5
    Green.pix <- Image.hue > 0.188 & Image.hue < 0.4167 & LeafMask > 0.5
    
    ###Label leaves and cycle through measurements
    print("Calculating standard leaf metrics....")
    LeafMask.lab <- bwlabel(LeafMask)
    Leaf.Results <- computeFeatures(LeafMask.lab,RawImage, xname = "Leaf",refnames=c("red","grn","blu"))
    LesionMask.lab <- LesionMask * LeafMask.lab
    Lesion.Results <- computeFeatures(LesionMask.lab,RawImage, xname = "Lesion",refnames=c("red","grn","blu"))
    
    if(dim(Lesion.Results)[1] != dim(Leaf.Results)[1]) {
      Leaf.Results <- Leaf.Results[rownames(Leaf.Results) %in% rownames(Lesion.Results),]  
    }
    
    Results <- cbind(Lesion.Results, Leaf.Results)
    rm(Lesion.Results, Leaf.Results)
    
    print("Calculating non-standard leaf metrics....")
    MoreResults <- as.data.frame(matrix(ncol=7))
    for(i in 1:max(LeafMask.lab)) {
      
      Lesion.Size <- sum(LesionMask.lab == i)
      Lesion.Grn <- sum(LesionMask.lab == i & Green.pix)
      Lesion.Ylw <- sum(LesionMask.lab == i & Yellow.pix)
      
      Leaf.Size <- sum(LeafMask.lab == i)
      Leaf.Grn <- sum(LeafMask.lab == i & Green.pix & !LesionMask)
      Leaf.Ylw <- sum(LeafMask.lab == i & Yellow.pix  & !LesionMask)
      
      Lesion.Prop <- Lesion.Size/Leaf.Size
      
      MoreResults[i,] <- c(Lesion.Size, Leaf.Size, Lesion.Prop, Leaf.Grn, Leaf.Ylw, Lesion.Grn, Lesion.Ylw)
      
    }
    
    colnames(MoreResults) <- c("Lesion.Size", "Leaf.Size", "Lesion.Prop", "Leaf.Grn", "Leaf.Ylw", "Lesion.Grn", "Lesion.Ylw")
    
    if(dim(Results)[1] != dim(MoreResults)[1]) {
      MoreResults <- MoreResults[rownames(MoreResults) %in% rownames(Results),]  
    }
    
    Results <- cbind(Results, MoreResults)
    
    
    ###Reorder Leaf Labels and Results
    print("Reordering leaf results for silly humans....")
    # find positions of objects and calculate distance matrix
    xy = computeFeatures.moment(LeafMask.lab)[, c('m.cx', 'm.cy')]
    Leaf.pos <- cbind(c(1,as.data.frame(xy)$m.cx), c(1,as.data.frame(xy)$m.cy))
    Temp.dist <- as.matrix(dist(Leaf.pos))
    #Gets the leaf closest to the top left and starts the key
    LeafOrder <- match(min(Temp.dist[-1,1]),Temp.dist[-1,1])
    #removes the reference top-left pixel
    Leaf.pos <- cbind(as.data.frame(xy)$m.cx, as.data.frame(xy)$m.cy)
    #recalculates distance matrix for objects
    Temp.dist <- as.matrix(dist(Leaf.pos))
    
    RowKey <- 0
    #loop through objects to reorder
    for(i in 1:dim(Temp.dist)[2]) {
      
      #Finds the ratio of x to y distance
        #if high, then objects are in the same line.  If low, objects are in diff lines.
      x.dif <- Leaf.pos[,1]-Leaf.pos[LeafOrder[i],1]
      y.dif <- Leaf.pos[,2]-Leaf.pos[LeafOrder[i],2]
      xy.ratio <- abs(x.dif/y.dif)
      
        # Tests if it has finished ordering all leaf objects
      if(length(LeafOrder) == dim(Leaf.pos)[1]) {
        break
        # Test if Image is done and missing values
      } else if(is.na(sum(xy.ratio > 1 & x.dif > 0) == 0)) {
        break
        # Test if it is at the end of the row
      } else if(sum(xy.ratio > 1 & x.dif > 0) == 0) {
        
        FirstLeaf <- LeafOrder[length(LeafOrder)-RowKey]
        
        # Find distances and ratio for all Leaf Objects relative to reference leaf 'FirstLeaf'.
        x.dif <- Leaf.pos[,1]-Leaf.pos[FirstLeaf,1]
        y.dif <- Leaf.pos[,2]-Leaf.pos[FirstLeaf,2]
        xy.ratio <- abs(x.dif/y.dif)
        
        # Find the leaf object with the minimum distance from the reference leaf object that is in the same row.
        Leaf <- match(min((Temp.dist[,FirstLeaf][xy.ratio < 1]), na.rm = TRUE), Temp.dist[,FirstLeaf])
        LeafOrder <- append(LeafOrder,Leaf)
        
        RowKey <- 0
        
        # If in the same row and 
      } else {
        Leaf <- match(min(Temp.dist[,LeafOrder[i]][xy.ratio > 1 & x.dif>0],na.rm = TRUE), Temp.dist[,LeafOrder[i]])
        LeafOrder <- append(LeafOrder,Leaf)
        
        RowKey <- RowKey+1
      }

    }
    
    LeafOrder <- unique(LeafOrder[!is.na(LeafOrder)])
    NewOrder <- 1:dim(xy)[1]
    MissingVal <- NewOrder[!NewOrder %in% LeafOrder]
    LeafOrder <- c(LeafOrder,MissingVal)
    Key <- cbind(LeafOrder,NewOrder)
    
    print("Reordering leaf mask objects....")
    for(i in 1:dim(Key)[2]){
      LeafMask.lab[LeafMask.lab == Key[i,1]] <- Key[i,2]  
      
    }
    
    ###Print results for image....
    
    print("Printing results....")
    Filtered <- Key[,1] %in% as.numeric(rownames(Results))
    Filtered2 <- match(LeafOrder[Filtered], rownames(Results))
    Results <- Results[Filtered2,]
    rownames(Results) <- Key[,2][Filtered]
    Results <- cbind(RawFiles[f], Results)
    colnames(Results)[1] <- "Image"
    write.csv(Results, paste(gsub(".jpg","_Results.csv",RawFiles[f])))
    
    print("Making reference image....")
        
        # get labels 
          Results <- data.frame("Object" = rownames(Results),Results)
        AddLabelOnPicture(Results,RawFiles[f])
        
  }
  print("All done. This had better go into Science or Nature for all my hard work!")

    }

```

# Contributions
Jason A Corwin (orcid 0000-0001-6455-8474) wrote the original code. Additional code improvement and corrections were contributed by Wei Zhang (orcid 0000-0002-5092-643X) and Celine Caseys (orcid 0000-0003-4187-9018). Funding was attributed to Daniel J Kliebenstein (orcid 0000-0001-5759-3175).


